\documentclass[addpoints]{exam}

\usepackage{amsmath,amssymb, amsthm}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\usepackage{forest}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{titling}
\usepackage{wasysym}
\usepackage{mdframed}
\usepackage{arabtex}
\usepackage{utf8}
\setcode{utf8}


% Header and footer.
\pagestyle{headandfoot}
\runningheadrule
\runningfootrule
\runningheader{CS 212, Fall 2024}{HW 2: Pumping Lemma, Context Free Languages}{Fall 2024}
\runningfooter{}{Page \thepage\ of \numpages}{}
\firstpageheader{}{}{}


\newcommand{\ghalib}{\text{\<غالب >}}
\newcommand{\mir}{\text{\<میر >}}
\newcommand{\U}{\text{\<الٹا >}}

\boxedpoints
% \printanswers %uncomment this

\title{Homework 2: Pumping Lemma, Context Free Languages}
\author{CS 212 Nature of Computation\\Habib University}
\date{Fall 2024}

\begin{document}
\maketitle

\begin{questions}

    \question Edsger W. Dijkstra once famously said \emph{``Besides a mathematical inclination, an exceptionally good mastery of one's native tongue is the most vital asset of a competent programmer.''}
    Inspired by this your lazy RA changes the terminal in the grammar given in your Nature of computation Homework.
    Consider the grammar containing the following productions.
    \begin{align*} 
      S \to &\; X\ghalib Y\\
      X \to &\; \ghalib X \mid \epsilon\\
      Y \to &\;  \ghalib Y \mid \mir Y \mid \epsilon
    \end{align*}

    \begin{parts}
    \part[10] Give a formal definition of the language generated by this grammar in set builder notation using regular expressions.
    
    % Use \ghalib to write غالب and Use \mir to write میر 
    \begin{solution}
        % Enter solution here 
    \end{solution}

    \part[10] Draw the parse tree under this grammar for the string: \ghalib\ghalib\ghalib\mir\ghalib.\footnote{ See \href{https://tex.stackexchange.com/a/214657/44301}{this post} on using the \texttt{forest} package to draw a parse tree.}
    \begin{solution}
        % Enter solution here
    \end{solution}

    \part[10] Argue whether this grammar is ambiguous.
    \begin{solution}
        % Enter solution here
    \end{solution}
    \end{parts}

    \question[15] Context free grammars are often used in compilers for syntax analysis. Your pretentious friend from a high-concept Christopher Nolan movie is making a compiler for a programming language. Now due to some time traveling nonsense, everything your friend speaks is in reverse. You believe that you can use this fact when making a context free grammar for syntax analysis for the programming language. You believe you just need to reverse an already existing programming language to get the job done.

    Given a set of alphabets, $\Sigma$, consider the operation, $\U:\Sigma^*\to\Sigma^*$, defined as follows.
    $$\U(w) = w_n\circ w_{n-1}\circ w_{n-2}\circ\ldots w_{1}, \text{ where } w = w_1\circ w_2\circ w_3\circ\ldots\circ w_n.$$
    $\U$ is extended to apply to a given language, $L$, as follows.
    $$\U(L) = \{ \U(w) \mid w\in L \}.$$
    Prove or disprove the claim that the class of context-free languages is closed under $\U$.

    % Use \U to write الٹا 
    \begin{solution}
        % Enter solution here
    \end{solution}


  
    \question Your friend Sud is in a Tom and Jerry-esque situation with a rat. Where he tries to catch the rat and the rat keeps outsmarting him. To taunt Sud, every time the rat runs out to steal some cheese it yells out its catchphrase which is of form ``\emph{RRRRRAAAAATTTT}'' given by the RegEx $R^iA^jT^k$ for $i,j,k \in \mathbb{Z}^+$. Now Sud, in a very Wile e Coyote-like fashion makes a contraption to catch the rat. The contraption has a DFA that recognizes the rat's catch phrase in order to detect the rat running around. The rat being much smarter than Sud decides to modify its catchphrase slightly such that such that it's catchphrase is no longer regular, so its not detected by Sud's machine. The new catch phrase is defined as $R^iA^jT^k$ where $i,j,k \in \mathbb{Z}^+$ and $i \neq 1$ or $j = k$.
    \begin{parts}
        \part[20] The rat graduated Habib University in 2021 with a CS degree, back in those days Nature of Computation was not a mandatory course. Due to this the rat doesn't know the pumping lemma. 
        Without using pumping lemma show that the following language $L_R$ is not regular:
        $L_R = \{R^iA^jT^k|\; i,j,k \in \mathbb{Z}^+ \text{ and } i \neq 1 \text{ or } j = k\}$.
        \begin{solution}
            % Enter solution here
        \end{solution}

        \part[15] Sud however has taken the Nature of Computation course, he knows that the set of context free languages is a superset of regular languages. He conjectures that the new language the rat is using to taunt him might be context free and he may be able to make a \href{https://en.wikipedia.org/wiki/Pushdown_automaton}{Push Down Automata (PDA)} for it. Prove or disprove the claim that the language, 
        $L_R = \{R^iA^jT^k|\; i,j,k \in \mathbb{Z}^+ \text{ and } i \neq 1 \text{ or } j = k\}$, is context-free.
        \begin{solution}
            % Enter solution here
        \end{solution}

    \end{parts}



    \question[20] Your friend Khubaib and Munawwar wish to evade taxes. Khubaib and Munawwar have decided to come up with a secret language to have all their embezzlement related correspondence in. As the tax auditor, uses a DFA to check for any suspicious activities, the language Khubaib and Munawwar come up with should not be regular. You were able to help them by giving them a language that you believe is not regular. You define the langauge as follows: $L_{\text{tax evasion}} = \{ w \in \{1, \smiley \}^* \mid w = w_1\smiley w_2\smiley \dots \smiley w_k \text{ for } k \geq 0, \text{ each } w_i \in 1^*, \text{ and } w_i \neq w_j \text{ for } i \neq j \}$. Prove or disprove that $L_{\text{tax evasion}}$ is not regular.

    % Use \smiley to write smile face emoji
    \begin{solution}
        % Enter solution here
    \end{solution}

\end{questions}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
